---
title: "Analyze str daily"
author: "Tourism Economics"
date: "April 14, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}

library(arlodr)
library(dplyr)
library(knitr)
library(readr)
library(readxl)
library(tidyr)
library(xts, warn.conflicts=FALSE)
library(purrr)
library(lubridate)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE)
```


```{r}

fpath <- c("~/Project/R projects/sf_2016/") 

```

```{r}

load(file=paste0(fpath,"output_data/str_daily.Rdata"))

```

Create a simple set of dates and days of week
```{r}

date <- seq(as.Date("1970-01-01"), as.Date("2030-01-01"), by = 1)
df <- data.frame(date)

date_df <- df %>%
  # add columns of month day
  mutate(month = lubridate::month(date, label=TRUE, abbr=TRUE)) %>%
  mutate(monthnum = lubridate::month(date, label=FALSE)) %>%
  mutate(day = day(date)) %>%
  mutate(wday = wday(date, label=TRUE)) %>%
  mutate(year = lubridate::year(date))

```

Join the simple date info onto the STR data
```{r}

str_daily_1 <- str_daily %>%
  left_join(date_df, by="date")  

```


Create seasonal categories
```{r}

seascat <- str_daily_1 %>%
  distinct(date) %>%
  # seasonal categories
  mutate(season1 = 
    ifelse(month %in% c("Jan", "Dec"), "01_Jan-Dec",
    ifelse(month %in% c("Feb", "Mar", "Apr", "May"), "02_Feb-Mar-Apr-May",
    ifelse(month %in% c("Jun"), "03_Jun",
    ifelse(month %in% c("Jul", "Aug", "Sep"), "04_Jul-Aug-Sep",
    ifelse(month %in% c("Oct"), "05_Oct",
    ifelse(month %in% c("Nov"), "06_Nov",
                            "other"))))))) %>%
  #year groupings
  mutate(yeargp = ifelse(year %in% c(2011, 2012), "yr11-12", 
                  ifelse(year %in% c(2013), "yr13",
                  ifelse(year %in% c(2014, 2015), "yr14-15",
                  "other")))) %>%
  select(date, season1, yeargp)
  
str_daily_2 <- str_daily_1 %>%
  left_join(seascat, by=c("date"))
  
```


How much data do we have for each geography in the total segment?
```{r}

str_daily_2 %>%
  filter(seg == "total") %>%
  group_by(geo_ttl, year) %>%
  summarize(a =n()) %>%
  ungroup() %>%
  spread(geo_ttl, a) %>%
   kable(digits=0, format.args = list(big.mark = ","))

```


Define some occupancy categories
```{r}
       
# occupancy categories
sfocccat1 <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "total") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31")) %>%
  mutate(sfocccat1 = cut(occ, c(0,.7,.8,.9,.95,1))) %>%
  select(date, sfocccat1)

# typical sf occupancy by season and wday
typocc <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "total") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31")) %>%
  group_by(geo_ttl, season1, wday) %>%
  summarize(mocc1 = mean(occ)) %>%
  ungroup() %>%
  select(-geo_ttl)

# work with the segment data
typocc_group <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "group") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31")) %>%
  filter(!(is.na(occ))) %>%
  group_by(season1, wday) %>%
  summarize(mgroup1 = mean(occ)) %>%
  ungroup() %>%
  select(season1, wday, mgroup1)

# sf group 
sfgroupcat1 <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "group") %>%
  filter(date >= as.Date("2012-01-01") & date <= as.Date("2015-12-31")) %>%
  mutate(sfgroupcat1 = cut(occ, c(0,.2,.4,.5,1))) %>%
  select(date, sfgroupcat1)


```

Join on these various new columns
```{r}

str_daily_3 <- str_daily_2 %>%
  left_join(sfocccat1, by=c("date")) %>%
  left_join(typocc, by=c("season1", "wday")) %>%
  left_join(typocc_group, by=c("season1", "wday")) %>%
  left_join(sfgroupcat1, by=c("date"))
  




```


# Data limitations
Looks like for SF City, we could do full year comparisons from 2011 to 2015. When we do comparisons with other geographies, we should do 2012 to 2015.  
The segmentation data for SF City is missing from 1/10/14 to 9/30/14.  

```{r}

dsfcity_seg <- str_daily_3 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31"))

dsfcity_1 <- str_daily_3 %>%
  filter(seg == "total") %>%
  filter(geo_ttl == "sfcity") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31"))

doaksf_1 <- str_daily_3 %>%
  filter(seg == "total") %>%
  filter(geo_ttl %in% c("sfcity", "oak")) %>%
  filter(date >= as.Date("2012-01-01") & date <= as.Date("2015-12-31"))

doaksf_seg <- str_daily_3 %>%
  filter(geo_ttl %in% c("sfcity", "oak")) %>%
  filter(date >= as.Date("2012-01-01") & date <= as.Date("2015-12-31"))

doaksf_seg_1 <- doaksf_seg %>%
  # drop the two dates in July that are missing from the SF data
  filter(!(date == as.Date("2015-07-26"))) %>%
  filter(!(date == as.Date("2015-07-27")))

```



## Frequency: Occupancy categories

Frequency of occupancy categories overall
```{r}
dsfcity_1 %>%
  group_by(sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by year
```{r}
dsfcity_1 %>%
  group_by(year, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by month (and number of days)
```{r}

dsfcity_1 %>%
  group_by(month, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  arrange(month) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(month, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # fill NAs with 0
  spread(sfocccat1, n, fill=0) %>%
  arrange(month) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by season1 (and number of days)
```{r}

dsfcity_1 %>%
  group_by(season1, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(season1, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # fill NAs with 0
  spread(sfocccat1, n, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by year and season1 (and number of days)
```{r}

dsfcity_1 %>%
  group_by(season1, year, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(season1, year, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # fill NAs with 0
  spread(sfocccat1, n, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by month, day of week (and number of days)
```{r}

dsfcity_1 %>%
  group_by(month, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # spread, but replace NA with 0
  spread(wday, freq, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(month, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # spread, but replace NA with 0
  spread(wday, n, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by season1, year group, day of week (and number of days)
```{r}

dsfcity_1 %>%
  group_by(season1, yeargp, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # spread, but replace NA with 0
  spread(wday, freq, fill=0) %>%
  kable(digits=c(0,0,0,rep(3,7)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(season1, yeargp, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # spread, but replace NA with 0
  spread(wday, n, fill=0) %>%
  kable(digits=c(0,0,0,rep(3,7)), format.args = list(big.mark = ","))

```


## Average occupancy by month, etc.

Average occupancy by month, wday
```{r}

dsfcity_1 %>%
  group_by(month, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

Average occupancy by season1, day of week
```{r}

dsfcity_1 %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))


```

Average occupancy by season1, year group, day of week
```{r}

dsfcity_1 %>%
  group_by(season1, yeargp, wday, sfocccat1) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))


```

## Occ and ADR

```{r}

dsfcity_1 %>%
  select(year, occ, adr) %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x=occ, y=adr, color=year)) +
  geom_point()


```

# Surrounding areas

## Initial explorations
```{r}

doaksf_1 %>%
  select(geo_ttl, occ, date) %>%
  spread(geo_ttl, occ) %>%
  ggplot(aes(x=sfcity, y=oak)) +
  geom_point()

```

With segment info 
```{r}

doaksf_seg_1 %>%
  select(geo_ttl, seg, occ, date) %>%
  unite(geoseg, geo_ttl, seg, sep="_", remove=TRUE) %>%
  spread(geoseg, occ) %>%
  ggplot(aes(x=sfcity_group, y=oak_total)) +
  geom_point()

```

## Oak daily as related to SF
Remember, SF segmentation data is missing from 1/10/14 to 9/30/14.  

### Occupancy grouped by SF-based categories
Take a look at what Oak occupancy looks like grouped by some of these SF-based categories.  

******
Grouped by SF occupancy category  

Overall (so what is oak occupancy, without regard to SF occupancy)
```{r}

gpoak_a <- doaksf_seg_1 %>%
  filter(geo_ttl == "oak") %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  mutate(sfocccat1 = "allcats") %>%
  select(sfocccat1, everything())

gpoak_a %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

```{r}

gpoak_b <- doaksf_seg_1 %>%
  filter(geo_ttl == "oak") %>%
  group_by(sfocccat1, season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) 

gpoak_b %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

Combine the tables
```{r}

gpoak_c <- bind_rows(gpoak_a, gpoak_b)

gpoak_c %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

```{r}

gpoak_d <- gpoak_c %>%
  group_by(season1) %>%
  mutate_each(funs(value = ./.[sfocccat1 == "allcats"]), -sfocccat1, -season1) %>%
  ungroup()

# calculate a mean of the weekdays across each row, removing NAs
gpoak_e <- gpoak_d %>% 
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE))

gpoak_e %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

******
Grouped by SF group occupancy contribution

Overall (so what is oak occupancy, without regard to SF occupancy)
```{r}

gpoak_a_2 <- doaksf_seg_1 %>%
  filter(geo_ttl == "oak") %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  mutate(sfgroupcat1 = "allcats") %>%
  select(sfgroupcat1, everything())

gpoak_a_2 %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

```{r}

gpoak_b_2 <- doaksf_seg_1 %>%
  filter(geo_ttl == "oak") %>%
  group_by(sfgroupcat1, season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) 

gpoak_b_2 %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

Combine the tables
```{r}

gpoak_c_2 <- bind_rows(gpoak_a_2, gpoak_b_2)

gpoak_c_2 %>%
  kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

```{r}
# following doesn't work because the Jun season (and maybe others) are missing from
# at least one of the category groupings

# gpoak_d_2 <- gpoak_c_2 %>%
#   group_by(season1) %>%
#   mutate_each(funs(value = ./.[sfgroupcat1 == "allcats"]), -sfgroupcat1, -season1) %>%
#   ungroup()
# 
# # calculate a mean of the weekdays across each row, removing NAs
# gpoak_e_2 <- gpoak_d_2 %>% 
#   mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE))
# 
# gpoak_e_2 %>%
#   kable(digits=c(1,rep(3,7)), format.args = list(big.mark = ","))

```

### Occupancy grouped by SF-based categories - modified approach
After discussing with Anthony, the suggestion 

