---
title: "Analyze str daily"
author: "Tourism Economics"
date: "April 14, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}

library(arlodr)
library(dplyr)
library(knitr)
library(readr)
library(readxl)
library(tidyr)
library(xts, warn.conflicts=FALSE)
library(purrr)
library(lubridate)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}

fpath <- c("~/Project/R projects/sf_2016/") 

```

```{r}

load(file=paste0(fpath,"output_data/str_daily.Rdata"))

```

# I. Set up data  

Create a simple set of dates and days of week  
```{r}

date <- seq(as.Date("1970-01-01"), as.Date("2030-01-01"), by = 1)
df <- data.frame(date)

date_df <- df %>%
  # add columns of month day
  mutate(month = lubridate::month(date, label=TRUE, abbr=TRUE)) %>%
  mutate(monthnum = lubridate::month(date, label=FALSE)) %>%
  mutate(day = day(date)) %>%
  mutate(wday = wday(date, label=TRUE)) %>%
  mutate(year = lubridate::year(date))

```

Join the simple date info onto the STR data  
```{r}

str_daily_1 <- str_daily %>%
  left_join(date_df, by="date")  

```

Create seasonal categories
```{r}

seascat <- str_daily_1 %>%
  distinct(date) %>%
  # seasonal categories
  mutate(season1 = 
    ifelse(month %in% c("Jan", "Dec"), "01_Jan-Dec",
    ifelse(month %in% c("Feb", "Mar", "Apr", "May"), "02_Feb-Mar-Apr-May",
    ifelse(month %in% c("Jun"), "03_Jun",
    ifelse(month %in% c("Jul", "Aug", "Sep"), "04_Jul-Aug-Sep",
    ifelse(month %in% c("Oct"), "05_Oct",
    ifelse(month %in% c("Nov"), "06_Nov",
                            "other"))))))) %>%
  #year groupings
  mutate(yeargp = ifelse(year %in% c(2011, 2012), "yr11-12", 
                  ifelse(year %in% c(2013), "yr13",
                  ifelse(year %in% c(2014, 2015), "yr14-15",
                  "other")))) %>%
  select(date, season1, yeargp)
  
str_daily_2 <- str_daily_1 %>%
  left_join(seascat, by=c("date"))
  
```

How much data do we have for each geography in the total segment?
```{r}

str_daily_2 %>%
  filter(seg == "total") %>%
  group_by(geo_ttl, year) %>%
  summarize(a =n()) %>%
  ungroup() %>%
  spread(geo_ttl, a) %>%
   kable(digits=0, format.args = list(big.mark = ","))

```

### Define some categories
Define some occupancy categories
```{r}
       
# occupancy categories
sfocccat1 <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "total") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31")) %>%
  mutate(sfocccat1 = cut(occ, c(0,.7,.8,.9,.95,1))) %>%
  select(date, sfocccat1)

#######
#
# typical occupancy by season and wday

# sf
typocc <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "total") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31")) %>%
  group_by(geo_ttl, season1, wday) %>%
  summarize(sfmocc1 = mean(occ),
            sfsdocc1 = sd(occ),
            sfnocc1 = n()) %>%
  ungroup() %>%
  select(-geo_ttl)

# oak
typocc_oak <- str_daily_2 %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31")) %>%
  group_by(geo_ttl, season1, wday) %>%
  summarize(oakmocc1 = mean(occ),
            oaksdocc1 = sd(occ),
            oaknocc1 = n()) %>%
  ungroup() %>%
  select(-geo_ttl)

########
#
# typical group occupancy

# sf
typocc_group <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "group") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31")) %>%
  filter(!(is.na(occ))) %>%
  group_by(season1, wday) %>%
  summarize(sfmgroup1 = mean(occ),
            sfsdgroup1 = sd(occ)) %>%
  ungroup() %>%
  select(season1, wday, sfmgroup1, sfsdgroup1)

########
#
# group categories

# sf 
sfgroupcat1 <- str_daily_2 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "group") %>%
  filter(date >= as.Date("2012-01-01") & date <= as.Date("2015-12-31")) %>%
  mutate(sfgroupcat1 = cut(occ, c(0,.2,.4,.5,1))) %>%
  select(date, sfgroupcat1)


```

Join on these various new columns
```{r}

str_daily_3 <- str_daily_2 %>%
  left_join(sfocccat1, by=c("date")) %>%
  left_join(typocc, by=c("season1", "wday")) %>%
  left_join(typocc_oak, by=c("season1", "wday")) %>%
  left_join(typocc_group, by=c("season1", "wday")) %>%
  left_join(sfgroupcat1, by=c("date"))

```

Using some of these new columns, calculate some additional measures to join
```{r}

# how does total occupancy for sf compare to its season1-wday mean, 
# output a list of dates with the results for those days

tempsf_a <- str_daily_3 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "total") %>%
  mutate(sfzocc1 = (occ-sfmocc1)/sfsdocc1) %>%
  select(date, sfzocc1)

# similar but for group occupancy
tempsf_b <- str_daily_3 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "group") %>%
  mutate(sfzgroup1 = (occ-sfmgroup1)/sfsdgroup1) %>%
  select(date, sfzgroup1)

# similar but for oak
tempoak_a <- str_daily_3 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(seg == "total") %>%
  mutate(oakzocc1 = (occ-oakmocc1)/oaksdocc1) %>%
  select(date, oakzocc1)

# join them on
str_daily_4 <- str_daily_3 %>%
  left_join(tempsf_a, by=c("date")) %>%
  left_join(tempsf_b, by=c("date")) %>%
  left_join(tempoak_a, by=c("date")) %>%
  # calculate a few categories
  mutate(sfzocc1_cat = cut(sfzocc1, c(-5,-2,-1,0,1,2,2.5))) %>%
  mutate(sfzgroup1_cat = cut(sfzgroup1, c(-5,-2,-1,0,1,2,3,5)))

```



### Data limitations
Looks like for SF City, we could do full year comparisons from 2011 to 2015. When we do comparisons with other geographies, we should do 2012 to 2015.  
The segmentation data for SF City is missing from 1/10/14 to 9/30/14.  

```{r}

dsfcity_seg <- str_daily_4 %>%
  filter(geo_ttl == "sfcity") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31"))

dsfcity_1 <- str_daily_4 %>%
  filter(seg == "total") %>%
  filter(geo_ttl == "sfcity") %>%
  filter(date >= as.Date("2011-01-01") & date <= as.Date("2015-12-31"))

doaksf_1 <- str_daily_4 %>%
  filter(seg == "total") %>%
  filter(geo_ttl %in% c("sfcity", "oak")) %>%
  filter(date >= as.Date("2012-01-01") & date <= as.Date("2015-12-31"))

doaksf_seg <- str_daily_4 %>%
  filter(geo_ttl %in% c("sfcity", "oak")) %>%
  filter(date >= as.Date("2012-01-01") & date <= as.Date("2015-12-31"))

# doaksf_seg_1 <- doaksf_seg %>%
#   # drop the two dates in July that are missing from the SF data
#   filter(!(date == as.Date("2015-07-26"))) %>%
#   filter(!(date == as.Date("2015-07-27")))

```

What do we have in dsfcity_1?
```{r}
nrow(dsfcity_1)
min(dsfcity_1$date)
max(dsfcity_1$date) 

# is occ blank for any rows?
dsfcity_1 %>%
  filter(is.na(occ)) %>%
  nrow()
  
```

What do we have in doaksf_seg for Oakland?
```{r}

a <- doaksf_seg %>%
  filter(seg == "total") %>%
  filter(geo_ttl == "oak")

nrow(a)
min(a$date)
max(a$date) 

# is occ blank for any rows?
a %>%
  filter(is.na(occ)) %>%
  nrow()
  
```

# II. Frequency: Occupancy categories

Frequency of occupancy categories overall
```{r}
dsfcity_1 %>%
  group_by(sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by year
```{r}
dsfcity_1 %>%
  group_by(year, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by wday
```{r}
dsfcity_1 %>%
  group_by(wday, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by month (and number of days)
```{r}

dsfcity_1 %>%
  group_by(month, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  arrange(month) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(month, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # fill NAs with 0
  spread(sfocccat1, n, fill=0) %>%
  arrange(month) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by season1 (and number of days)
```{r}

dsfcity_1 %>%
  group_by(season1, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(season1, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # fill NAs with 0
  spread(sfocccat1, n, fill=0) %>%
  kable(digits=c(1,rep(3,5)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by year and season1 (and number of days)
```{r}

dsfcity_1 %>%
  group_by(season1, year, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # fill NAs with 0
  spread(sfocccat1, freq, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(season1, year, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # fill NAs with 0
  spread(sfocccat1, n, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by month, day of week (and number of days)
```{r}

dsfcity_1 %>%
  group_by(month, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  # spread, but replace NA with 0
  spread(wday, freq, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(month, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  # spread, but replace NA with 0
  spread(wday, n, fill=0) %>%
  kable(digits=c(1,rep(3,6)), format.args = list(big.mark = ","))

```

Frequency of occupancy categories by season1, year group, day of week (and number of days)
```{r}

dsfcity_1 %>%
  group_by(season1, yeargp, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  complete(season1, yeargp, wday, sfocccat1) %>%
  # spread, but replace NA with 0
  spread(wday, freq, fill=0) %>%
  kable(digits=c(0,0,0,rep(3,9)), format.args = list(big.mark = ","))

dsfcity_1 %>%
  group_by(season1, yeargp, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  ungroup() %>%
  complete(season1, yeargp, wday, sfocccat1) %>%
  # spread, but replace NA with 0
  spread(wday, n, fill=0) %>%
  kable(digits=c(0,0,0,rep(3,9)), format.args = list(big.mark = ","))

```

# III.A. Average occupancy by month, etc.

Average occupancy by month, wday
```{r}

dsfcity_1 %>%
  group_by(month, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```

Average occupancy by season1, day of week
```{r}

dsfcity_1 %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))


```

Average occupancy by season1, year group, day of week
```{r}

dsfcity_1 %>%
  group_by(season1, yeargp, wday, sfocccat1) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  complete(season1, yeargp, wday, sfocccat1) %>%
  spread(wday, occ) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))


```

Average ADR by season1, year group, day of week
```{r}

dsfcity_1 %>%
  group_by(season1, yeargp, wday, sfocccat1) %>%
  summarise (adr = mean(adr)) %>%
  ungroup() %>%
  complete(season1, yeargp, wday, sfocccat1) %>%
  spread(wday, adr) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```



Average ADR change by occupancy category
(So this is the average ADR by year and occ category, and then the percentage change relative to the yr11-12 period as a base.)
```{r}

dsfcity_1 %>%
  group_by(yeargp, sfocccat1) %>%
  summarise (adr = mean(adr)) %>%
  ungroup() %>%
  complete(yeargp, sfocccat1) %>%
  mutate(adr_chg = adr / adr[yeargp == "yr11-12"] -1) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```



## Occ and ADR

```{r}

dsfcity_1 %>%
  select(year, occ, adr) %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x=occ, y=adr, color=year)) +
  geom_point()


```

# IV. Oak daily as related to SF

Remember, SF segmentation data is missing from 1/10/14 to 9/30/14.   

Initial explorations  
```{r}

doaksf_1 %>%
  select(geo_ttl, occ, date) %>%
  spread(geo_ttl, occ) %>%
  ggplot(aes(x=sfcity, y=oak)) +
  geom_point()

```

With segment info   
```{r}

doaksf_seg %>%
  select(geo_ttl, seg, occ, date) %>%
  unite(geoseg, geo_ttl, seg, sep="_", remove=TRUE) %>%
  spread(geoseg, occ) %>%
  ggplot(aes(x=sfcity_group, y=oak_total)) +
  geom_point()

```

## IV.A. Occupancy impact to Oak grouped by SF-based categories: _Initial approach_
This initial approach is based on the simple occupancy categories for SF, rather than z-scores)  
  
******
### Grouped by SF occupancy category
Take a look at what Oak occupancy looks like grouped by some of these SF-based categories.  

Overall (so what is oak occupancy, without regard to SF occupancy)
```{r}

gpoak_a <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  mutate(sfocccat1 = "allcats") %>%
  select(sfocccat1, everything())

gpoak_a %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

# mean by category
gpoak_b <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  group_by(sfocccat1, season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) 

gpoak_b %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

#  combine the tables
gpoak_c <- bind_rows(gpoak_a, gpoak_b) %>%
  tidyr::complete(sfocccat1, season1) %>%
  arrange(season1) 

gpoak_c %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

# report frequency in each category
gpoak_b_freq <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  filter(!(is.na(sfocccat1))) %>%
  group_by(season1, wday, sfocccat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  spread(wday, freq, fill=0) 

gpoak_b_freq %>%
  select(sfocccat1, everything()) %>%
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE)) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```

Compare as ratio to mean for that season-wday
```{r}

gpoak_d <- gpoak_c %>%
  group_by(season1) %>%
  mutate_each(funs(value = ./.[sfocccat1 == "allcats"]), -sfocccat1, -season1) %>%
  ungroup()

# calculate a mean of the weekdays across each row, removing NAs
gpoak_e <- gpoak_d %>% 
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE))

gpoak_e %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```

******
### Grouped by SF group occupancy contribution

```{r}

# Overall (so what is oak occupancy, without regard to SF occupancy)
gpoak_a_2 <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(!(is.na(sfgroupcat1))) %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  mutate(sfgroupcat1 = "allcats") %>%
  select(sfgroupcat1, everything())

gpoak_a_2 %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

# mean by categories
gpoak_b_2 <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(!(is.na(sfgroupcat1))) %>%
  group_by(sfgroupcat1, season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) 

gpoak_b_2 %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

# combine the tables
gpoak_c_2 <- bind_rows(gpoak_a_2, gpoak_b_2) %>%
  # complete fills in the missing combinations because otherwise
  # there are some categories that are missing (dropped in the summarize step)
  tidyr::complete(sfgroupcat1, season1) %>%
  arrange(season1) 

gpoak_c_2 %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

# report frequency in each category
gpoak_b_2_freq <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  filter(!(is.na(sfgroupcat1))) %>%
  group_by(season1, wday, sfgroupcat1) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  spread(wday, freq, fill=0) 

gpoak_b_2_freq %>%
  select(sfgroupcat1, everything()) %>%
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE)) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```

Compare as ratio to mean for that season-wday
```{r}

gpoak_d_2 <- gpoak_c_2 %>%
  group_by(season1) %>%
  mutate_each(funs(value = ./.[sfgroupcat1 == "allcats"]), -sfgroupcat1, -season1) %>%
  ungroup()

# calculate a mean of the weekdays across each row, removing NAs
gpoak_e_2 <- gpoak_d_2 %>%
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE))

gpoak_e_2 %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```

## IV.B. Occupancy impact to Oak grouped by SF-based categories: _Modified approach_
This initial approach is based on z-scores of various occupancy measures in SF

After discussing with Anthony, the suggestion was to look at days when SF occupancy is a standard deviation above its average, and whether Oakland is also

Graph of occupancy z-scores (number of standard deviations above or below mean)
```{r}

doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  ggplot(aes(x=sfzocc1, y=oakzocc1)) +
  geom_point()
  
```


******
### Grouped by SF total occupancy (using z-score)  
Oakland occupancy during various sf occupancy categories based on z-score of total occupancy
```{r}
# mean occupancy in all categories
tk2oak_a <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  mutate(sfzocc1_cat = "allcats") %>%
  select(sfzocc1_cat, everything())

tk2oak_a %>%
  kable(digits=c(1,rep(3,8)), format.args = list(big.mark = ","))

# mean occupancy by categories
tk2oak_b <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  group_by(season1, wday, sfzocc1_cat) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) 

tk2oak_b %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

# combine the tables
tk2oak_c <- bind_rows(tk2oak_a, tk2oak_b) %>%
  complete(sfzocc1_cat, season1) %>%
  arrange(season1)

tk2oak_c %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

# report frequency in each category
tk2oak_b_freq <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  filter(!(is.na(sfzocc1_cat))) %>%
  group_by(season1, wday, sfzocc1_cat) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  spread(wday, freq, fill=0) 

tk2oak_b_freq %>%
  select(sfzocc1_cat, everything()) %>%
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE)) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```

Compare as ratio to mean for that season-wday
```{r}

tk2oak_d <- tk2oak_c %>%
  group_by(season1) %>%
  mutate_each(funs(value = ./.[sfzocc1_cat == "allcats"]), -sfzocc1_cat, -season1) %>%
  ungroup()

# calculate a mean of the weekdays across each row, removing NAs
tk2oak_e <- tk2oak_d %>%
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE))

tk2oak_e %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```


Graph of occupancy z-scores (group for sf, total for oak). I think the striations come from the fact that we've calculated a z-score relative to 42 different category means for each. So dates that fit in a certain season-wday are being divided by the same number, and maybe the relationship between those two numbers (sf and oak) stays as a consistent underlying pattern).
```{r}

doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  ggplot(aes(x=sfzgroup1, y=oakzocc1)) +
  geom_point()
  
```


******
### Grouped by SF group occupancy (using z-score)  

Oakland occupancy during various sf group demand categories based on z-score of group occupancy. The NAs in sfzgroup1_cat are days when we don't have SF group data. I decided to drop them so that the relative frequency would make sense.
```{r}
# mean occupancy in all categories
tk2oak_a_2 <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  filter(!(is.na(sfzgroup1_cat))) %>%
  group_by(season1, wday) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) %>%
  mutate(sfzgroup1_cat = "allcats") %>%
  select(sfzgroup1_cat, everything())

tk2oak_a_2 %>%
  kable(digits=c(1,rep(3,8)), format.args = list(big.mark = ","))

# mean occupancy by categories
tk2oak_b_2 <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  filter(!(is.na(sfzgroup1_cat))) %>%
  group_by(season1, wday, sfzgroup1_cat) %>%
  summarise (occ = mean(occ)) %>%
  ungroup() %>%
  spread(wday, occ) 

tk2oak_b_2 %>%
  kable(digits=c(1,rep(3,8)), format.args = list(big.mark = ","))

# combine the tables
tk2oak_c_2 <- bind_rows(tk2oak_a_2, tk2oak_b_2) %>%
  complete(sfzgroup1_cat, season1) %>%
  arrange(season1)

tk2oak_c_2 %>%
  kable(digits=c(1,rep(3,8)), format.args = list(big.mark = ","))

# report frequency in each category
tk2oak_b_2_freq <- doaksf_seg %>%
  filter(geo_ttl == "oak") %>%
  filter(seg == "total") %>%
  filter(!(is.na(sfzgroup1_cat))) %>%
  group_by(season1, wday, sfzgroup1_cat) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  spread(wday, freq, fill=0) 

tk2oak_b_2_freq %>%
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE)) %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```

Compare as ratio to mean for that season-wday
```{r}

tk2oak_d_2 <- tk2oak_c_2 %>%
  group_by(season1) %>%
  mutate_each(funs(value = ./.[sfzgroup1_cat == "allcats"]), -sfzgroup1_cat, -season1) %>%
  ungroup()

# calculate a mean of the weekdays across each row, removing NAs
tk2oak_e_2 <- tk2oak_d_2 %>%
  mutate(mean=rowMeans(.[, sapply(., is.numeric)], na.rm=TRUE))

tk2oak_e_2 %>%
  kable(digits=c(1,rep(3,9)), format.args = list(big.mark = ","))

```
